Quantifying the ease of viral escape from broad and narrow antibodies to influenza hemagglutinin

### 2017-10-08 paper/manuscript.tex

\section*{INTRODUCTION}
Nearly all viruses exhibit some antigenic variation.
However, the extent of this variation ranges widely across viruses and antibodies.
Broadly neutralizing antibodies against influenza virus are of great interest... \cite{corti2017tackling}.

>>>>

\section*{INTRODUCTION}
Nearly all viruses show some antigenic variation.
However, the extent of this variation ranges widely.
For instance, although both measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic,drexler2014robustness} exhibit antigenic variation, the magnitude of this variation is small. 
Therefore, immunity elicited by infection or vaccination provides life-long protection against both these viruses~\citep{?}.
In contrast, some viruses exhibit much more extensive antigenic variation.
An example is human influenza virus.
Infection with influenza virus provides life-long immunity to the infecting strain~\citep{?}.
But influenza virus undergoes such rapid antigenic evolution that the effectiveness of this immunity to new viral strains is almost completely eroded within a few years~\citep{?}.

One possible explanation for the reason that viruses exhibit differences in the extent of antigenic variation is that they have disparate inherent evolutionary capacities to escape immunodominant antibodies~\citep{lipsitch2007patterns,cobey2014pathogen,fulton2015mutational}.
According to this explanation, human influenza virus undergoes rapid antigenic drift because most neutralizing antibodies target epitopes on the viral hemagglutinin protein that are highly tolerant of mutational change.
This explanation is supported by classic experiments showing that it is easy to select viral mutants that escape most anti-hemagglutinin antibodies~\citep{yewdell1979antigenic,webster1980determination}, as well as by the observation that mutations that alter antigenicity arise frequently during influenza's evolution at the global scale~\citep{koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction} and even within individual humans with long-term infections~\citep{xue2017parallel}.
A corollary of this explanation is that the virus's capacity for antigenic drift would be reduced if most antibodies instead targeted epitopes that were less mutationally tolerant.

Verifying this corollary has become of practical importance with the discovery of broadly neutralizing antibodies against influenza virus.
These antibodies typically target conserved epitopes in hemagglutinin's stalk~\citep{?} or receptor-binding pocket~\citep{?}, and neutralize a wide range of viral strains.
Broad antibodies usually make only a modest contribution to natural human immunity, since they are less abundant than antibodies to the antigenically variable head of hemagglutinin~\citep{?}.
However, major efforts are now underway to elicit broad antibodies by vaccination~\citep{?} or administer them directly as therapeutics or prophylactics~\citep{?}.

If these efforts succeed, the epitopes of broadly neutralizing antibodies will then be under strong antigenic selection in human influenza virus.
Certainly there is precedent for the idea that changing the immune status of the host population can alter influenza virus evolution: ...
Of course, it remains unclear if these epitopes are more conserved because they are inherent intolerant or just aren't under pressure.
For instance, non-human influenza doesn't undergo much antigenic drift.
Knowing is important because success of a vaccine would put them under pressure.

Broadly neutralizing antibodies against influenza virus are of great interest... \cite{corti2017tackling}.
